Appendix

### 2025-03-31 manuscript_rmarkdown/manuscript.Rmd
Added  simulation data to the main manuscript.

To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable),
we minimised $\sum_{i=1}^{m_d = 100}\tau_{n, \kappa = 0.70, i}$, i.e. the mean of residual 
errors of the central 70% of elements of each sequence $\mathbf{X}_i$, summed over all sequences.

@@

>>>>

To find the optimal values for the weighting function parameters $\delta_1$ and $\delta_2$ (if applicable),
we minimised $\sum_{i=1}^{m_d = 100}\tau_{n_i, \kappa = 0.70, i}$, i.e. the mean of residual 
errors of the central 70% of elements of each sequence $\mathbf{X}_i$, summed over all sequences.

@@

# Yeo-Johnson on clean and dirty data.
clean_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "clean")
dirty_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty")
shifted_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty_shifted")

data <- rbind(clean_data, dirty_data, shifted_data, fill = TRUE)
data$dataset <- factor(data$dataset, levels = c("clean", "dirty", "dirty_shifted"))

# Add rank based on residual error for central normality.
data[outlier == FALSE, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]
data[outlier == TRUE, "method_rank" := rank(residuals_central), by = c("experiment_id", "outlier", "dataset")]

table_data <- data[
  ,
  list(
    "avg." = round(mean(method_rank), digits = 2L),
    "10%" = quantile(method_rank, 0.10),
    "50%" = quantile(method_rank, 0.50),
    "90%" = quantile(method_rank, 0.90)
  ),
  by = c("transformation", "outlier", "dataset")
]

data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg.")

\begin{table}
\begin{center}
\caption{
Comparison of average rank between transformation methods based on either residual error (without outliers) or residual error of the central portion
(with outliers; $\kappa = 0.70$) over 3 datasets with 10000 sequences each. The clean dataset consists of sequences derived through inverse Yeo-Johnson transformation
of data sampled from a standard normal distribution. The dirty dataset contains sequences sampled from asymmetric generalised normal distributions, centred at 0.
The shifted dataset also contains sequences sampled from asymmetric generalised normal distributions, but centred at 100, and scaled by 0.001.
Several transformation methods include normalisation before transformation, indicated by z-score normalisation (norm.) or robust scaling.
A rank of 1 is the best and a rank of 9 the worst. For each dataset, the best ranking transformation is marked in bold.
}
\label{tab:comparison_methods_simulations}
\begin{tabular}{l | l r r r r r r}

\toprule
& dataset: & \multicolumn{2}{c}{clean} & \multicolumn{2}{c}{dirty} & \multicolumn{2}{c}{shifted} \\
transformation & outliers: & no & yes & no & yes & no & yes \\

\midrule

none                                  & &         8.45  &         7.52  &         7.29  &         5.99  &         7.39  &         6.39 \\
conventional                          & &         3.64  &         5.57  &         3.78  &         7.04  &         6.45  &         5.79 \\
conventional (z-score norm.)          & &         4.68  &         5.78  &         6.29  &         5.10  &         5.34  &         5.08 \\
conventional (robust scaling)         & &         4.00  &         6.74  &         5.28  &         5.78  &         4.32  &         5.79 \\
Raymaekers-Rousseeuw                  & &         4.41  &         3.66  &         3.92  &         4.40  &         6.23  &         5.50 \\
Raymaekers-Rousseeuw (z-score norm.)  & &         5.21  &         3.16  &         6.13  &         4.24  &         5.18  &         4.15 \\
Raymaekers-Rousseeuw (robust scaling) & &         4.62  &         3.42  &         5.22  &         4.38  &         4.26  &         4.30 \\
invariant                             & & \textbf{3.05} &         6.95  & \textbf{3.03} &         6.16  & \textbf{2.45} &         6.13 \\
robust invariant                      & &         6.95  & \textbf{2.20} &         4.07  & \textbf{1.90} &         3.38  & \textbf{1.87} \\

\bottomrule
\end{tabular}
\end{center}
\end{table}

### 2025-03-31 manuscript_rmarkdown/manuscript_appendix.Rmd
Added  simulation data to the main manuscript.

>>>>

# Yeo-Johnson on clean and dirty data.
clean_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "clean", method = "box_cox")
dirty_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty", method = "box_cox")
shifted_data <- .assess_simulated_data(manuscript_dir = manuscript_dir, data_type = "dirty_shifted", method = "box_cox")

data <- rbind(clean_data, dirty_data, shifted_data, fill = TRUE)
data$dataset <- factor(data$dataset, levels = c("clean", "dirty", "dirty_shifted"))

# Add rank based on residual error for central normality.
data[, "method_rank" := rank(residuals), by = c("experiment_id", "outlier", "dataset")]

table_data <- data[
  ,
  list(
    "avg." = round(mean(method_rank), digits = 2L),
    "10%" = quantile(method_rank, 0.10),
    "50%" = quantile(method_rank, 0.50),
    "90%" = quantile(method_rank, 0.90)
  ),
  by = c("transformation", "outlier", "dataset")
]

data.table::dcast(transformation ~ dataset + outlier, data = table_data, value.var = "avg.")
